Automated Example Oriented REST API Documentation

### 2017-09-24 introduction.tex
add changes based on feedback from craig + mea

\chapter{Introduction}
Application Programming Interfaces, commonly known as APIs, are used to express a software component in terms of its operations, inputs, outputs, and their types\footnote{\url{https://en.wikipedia.org/wiki/Application_programming_interface}}. Robillard defines an API as follows: An API is the interface to implement functionality that developers can access to perform various tasks \cite{Robillard_a_field_study, Robillard_what_makes}. APIs enable multiple software components to interact with each other.

REST APIs are a type of APIs that are used to integrate software components using web technologies. Fielding defined Representational State Transfer or REST as an architectural style for developing distributed hypermedia systems \cite{Fielding_rest}. For example, GitHub, an online code repository hosting tool, has a REST API that can be used by third-party applications to integrate with GitHub features. The GitHub REST API\footnote{\url{https://developer.github.com/v3/repos/\#create}} has a resource called \texttt{Repository} to denote a code repository that can be identified by the URL \url{/user/repos}. To create a new \texttt{Repository} for a user, the GitHub API can be invoked via HTTP $POST$ at \url{/user/repos} using the following JSON representation of a \texttt{Repository}:

@@

While there are overlaps between the documentation requirements of local and REST APIs, there are significant requirements that are unique to each. The documentation of most local APIs contain a structural definition of the API classes and methods. API usage examples for local APIs are documented using the same programming language of the API. On the other hand, REST APIs are defined using HTTP specific parameters such as API endpoints, query parameters, HTTP headers and body. REST API clients can use any programming language agnostic to the API implementation language. Due to such structural and implementation differences between local and REST APIs, the existing technique and tools for local API documentation cannot be readily used to generate REST API documentation.

@@

\begin{itemize}
  \item S1 - API developer prepares an example API with the required HTTP parameters and request body.
  \item S2 - API developer uses a REST API client to make the example API call.
  \item S3 - the API developer records the HTTP data.
  \item S4 - API developer then edits the recorded data so that only relevant content is selected for documentation.
  \item S5 - API developers use any custom content to describe the API example.
  \item S6 - API developers combine the custom content with the edited content from the HTTP traffic into HTML to publish to the web.
  \item S7 - API developer adds custom overview information to explain API concepts and general rules, and publishes the final documentation so other developers can learn the API
\end{itemize}

@@

\section{Research Goals}
The overarching goal of my research was to improve the current state of REST API documentation process by automating the repeating manual steps in the aforementioned workflow. More specifically, I aimed to solve the following research goals:

\begin{itemize}
  \item \textit{RG1: To identify a set of REST API documentation requirements.} Existing research on API usability has derived a number of API documentation requirements primarily focusing on local APIs. Due to the structural and implementation differences between local and REST APIs, I aimed to identity REST API specific documentation requirements in my research. This has several implications as follows: having access to a scientifically derived list of REST API documentation requirements allows REST API developers to make an informed decision related to planning and execution of their documentation efforts. The requirements provide a background research that can be used by practitioners to compare and contrast existing REST API documentation tools while selecting an appropriate tool for their API. REST API documentation tool developers can also use the requirements to build necessary features.
  \item \textit{RG2: To develop and evaluate a technique to overcome the commonly observed but unresolved challenges related to REST API documentation.} I aimed to devise a reusable technique so that REST APIs can be documented following the identified REST API documentation requirements without following manual or bespoke methods. It has the following implications: REST API developers can utilize the technique to overcome the common challenges with the generation and maintenance of their API documentation. REST API client developers can access a usable API documentation.
\end{itemize}

\section{Research Objectives}
Related to RG1 and RG2, I aimed to address the following research objectives:

\begin{itemize}
  \item \textit{RO1: How are evolving REST APIs documented?} This question is aimed to analyze the gap between current industry practice around the documentation of evolving REST APIs and the proposed solutions from existing literature. This objective helped me to identify the research opportunity to work on unresolved problems related to REST API documentation.
  \item \textit{RO2: What are the requirements for usable REST API documentation?} This objective provides a necessary background work for improving the state of REST API documentation process.
  \item \textit{RO3: What technique can be used to document REST APIs to meet the requirements?} This is the central research objective. I researched to develop a new technique that can fill the gap between existing REST API documentation methods and the identified requirements.
  \item \textit{RO4: Is the proposed technique feasible for REST API developers?} An evaluation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
  \item \textit{RO5: How does a documentation using the proposed technique impact REST API client developers?} This question investigates the impact of the proposed technique from the perspective of the REST API client developers.
\end{itemize}

\section{Cohesion and Unitary Report}
To address the aforementioned research goals and objectives, my research strategy included five sub-projects, result of which have been published as papers in peer-reviewed international venues. Table~\ref{table:chapter_summary} shows a high level overview of the flow of arguments between the sub-projects to meet the research goals.

\begin{table}
  \caption{Summary and Flow of Arguments of the Sub-projects}
  \begin{tabular}{|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{4.5cm}|p{6cm}|}
  \hline
  Chapter & Goal & Objective & Research Project & Primary Outcome \\
  \hline
  Chapter~\ref{chapter:case_study}  & RG1 & RO1 & A case study of Web API evolution \cite{sohan2015case}. & A set of challenging problems related to evolving Web APIs. Selected REST API documentation as the problem of interest for this research. \\
  \hline
  Chapter~\ref{chapter:spy_rest} & RG1 and RG2 & RO2 and RO3 & SpyREST: Automated RESTful API documentation using an HTTP proxy server \cite{sohan2015spyrest}. & A list of requirements and the concept of using interception to generate REST API documentation to satisfy the requirements. \\
  \hline
  Chapter~\ref{chapter:demo_paper} & RG2 & RO3 & SpyREST in action: An automated RESTful API documentation tool \cite{sohan2015spyrest_tool}. & A reusable tool and a demonstration of the proposed REST API documentation technique. \\
  \hline
  Chapter~\ref{chapter:cisco_study} & RG2 & RO4 & Automated example oriented REST API documentation at Cisco \cite{sohan_cisco}. & An empirical evidence that the proposed technique is feasible for practical use, and a list of problems for potential future research.\\
  \hline
  Chapter~\ref{chapter:controlled_study} & RG2 & RO5 & A study of the effectiveness of usage examples in REST API documentation \cite{sohan_vlhcc}. & An empirical evidence that REST API client developers can overcome common obstacles when the documentation contains usage examples, as suggested in the proposed technique.\\
  \hline
\end{tabular}
\label{table:chapter_summary}
\end{table}

@@

In Chapter~\ref{chapter:spy_rest}, I discuss my second research project where I derived a set of REST API documentation requirements and proposed a new technique to satisfy the requirements \cite{sohan2015spyrest}. Here, I present a list of REST API documentation requirements by studying the literature and applying the findings from Chapter~\ref{chapter:case_study}. I discuss the available REST API documentation techniques and compare against the requirements. I present, SpyREST, a novel technique to intercept example REST API calls using an HTTP proxy server to collect and automatically synthesize API traffic to generate and update API documentation to satisfy the identified requirements. The primary advantage of the presented technique is that the documentation is generated from executable code. As a result, a continuously updated REST API documentation with realistic usage examples is achievable with a reduced manual effort.

I present my third research project, related to RG2, in Chapter~\ref{chapter:demo_paper}, where I present a prototype implementation of the aforementioned technique \cite{sohan2015spyrest_tool}. In this prototype implementation, I discuss the process that transforms a number of example HTTP requests and responses into a usable REST API documentation. I also present an example where automated test code is used to generate and maintain a continuously updated REST API documentation using the developed tool.

>>>>

\chapter{Introduction}

Robillard defines an API as follows: An API is the interface to implement functionality that developers can access to perform various tasks \cite{Robillard_a_field_study, Robillard_what_makes}. Application Programming Interfaces, commonly known as APIs, are used to express a software component in terms of its operations, inputs, outputs, and their types\footnote{\url{https://en.wikipedia.org/wiki/Application_programming_interface}}. APIs enable multiple software components to interact with each other. Web APIs allow software-to-software interconnectivity using web technologies as the transport mechanism of information between systems.

REST APIs are a type of API that are used to integrate software components using web technologies. Fielding defined Representational State Transfer or REST as an architectural style for developing distributed hypermedia systems \cite{Fielding_rest}. He described the design of modern Web as an example of a REST-style architecture. REST-style architecture allows software components to satisfy the data transfer needs of the modern Internet-scale applications by levergading a generic interface based on URIs and Internet media types that allow intemediate processing such as caching to improve system performance.

REST APIs allow access to application resources over HTTP. While REST describes a generic protocol for accessing application resources, the applications are responsible for the specific representation of the resources. For example, a REST API may support multiple Internet media types such as JSON, XML, CSV for a single application resource. Similarly, two diffenent REST APIs may use the same media type but contain entirely different formats to represent their objects.

For example, GitHub, an online code repository hosting tool, has a REST API that can be used by third-party applications to integrate with GitHub features. The GitHub REST API\footnote{\url{https://developer.github.com/v3/repos/\#create}} has a resource called \texttt{Repository} to denote a code repository that can be identified by the URL \url{/user/repos}. To create a new \texttt{Repository} for a user, the GitHub API can be invoked via HTTP $POST$ at \url{/user/repos} using the following JSON representation of a \texttt{Repository}:

@@

While there are overlaps between the documentation requirements of local and REST APIs, there are significant requirements that are unique to each. The documentation of most local APIs contain a structural definition of the API classes and methods. API usage examples for local APIs are documented using the same programming language of the API. On the other hand, REST APIs are defined using HTTP specific parameters such as API endpoints, query parameters, HTTP headers and body. REST API clients can be written in any programming language agnostic to the API implementation language. Due to such structural and implementation differences between local and REST APIs, the existing technique and tools for local API documentation cannot be readily used to generate REST API documentation.

@@

\begin{itemize}
  \item S1 - API developer prepares an example API with the required HTTP parameters and request body.
  \item S2 - API developer uses a REST API client to make the example API call.
  \item S3 - the API developer records the HTTP data.
  \item S4 - API developer then edits the recorded data so that only relevant content is selected for documentation.
  \item S5 - API developers use any custom content to describe the API example.
  \item S6 - API developers combine the custom content with the edited content from the HTTP traffic into HTML to publish on the web.
  \item S7 - API developer adds custom overview information to explain API concepts and general rules, and publishes the final documentation so other developers can learn the API.
\end{itemize}

@@

\section{Research Goals}
The overarching goal of my research is to improve the current state of the proces of REST API documentation process by automating the repeating manual steps in the aforementioned workflow. More specifically, I aimed to solve the following research goals:

\begin{itemize}
  \item \textit{RG1: To identify a set of REST API documentation requirements.} Existing research on API usability has derived a number of API documentation requirements primarily focusing on local APIs. Due to the structural and implementation differences between local and REST APIs, I aimed to identity REST API specific documentation requirements in my research. This has several implications as follows: having access to a scientifically derived list of REST API documentation requirements allows REST API developers to make an informed decision related to planning and execution of their documentation efforts. The requirements provide a background research that can be used by practitioners to compare and contrast existing REST API documentation tools while selecting an appropriate tool for their API. REST API documentation tool developers can also use the requirements to build necessary features.
  \item \textit{RG2: To develop and evaluate a technique to overcome the commonly observed challenges related to REST API documentation.} I aimed to devise a reusable technique so that REST APIs can be documented following the identified REST API documentation requirements without following manual or bespoke methods. It has the following implications: REST API developers can utilize the technique to overcome the common challenges with the generation and maintenance of their API documentation, and REST API client developers can access an API documentation.
\end{itemize}

\section{Research Objectives}
Related to RG1 and RG2, I aimed to address the following research questions as my objectives:

\begin{itemize}
  \item \textit{RO1: How are evolving REST APIs documented?} This question is aimed to analyze the gap between current industry practice around the documentation of evolving REST APIs and the proposed solutions from existing literature. This objective helped me to identify the research opportunity to work on unresolved problems related to REST API documentation.
  \item \textit{RO2: What are the requirements for usable REST API documentation?} This objective provides a necessary background for improving the state of REST API documentation process.
  \item \textit{RO3: What technique can be used to document REST APIs to meet the requirements?} This is the central research objective. I researched to develop a new technique that can fill the gap between existing REST API documentation methods and the identified requirements.
  \item \textit{RO4: Is the proposed technique a practical one for REST API developers?} A qualitiative evaluation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
  \item \textit{RO5: How does a documentation using the proposed technique impact REST API client developers?} This question investigates the impact of the proposed technique from the perspective of the REST API client developer effort and success rate while performing API tasks.
\end{itemize}

\section{Cohesion and Unitary Report}
To address the aforementioned research goals and objectives, my research strategy included five sub-projects, have resulted in publications at peer-reviewed international conferences. Table~\ref{table:chapter_summary} shows a high level overview of the flow of arguments between the sub-projects to meet the research goals.

\begin{table}
  \caption{Summary and Flow of Arguments of the Sub-projects}
  \begin{tabular}{|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{4.5cm}|p{6cm}|}
  \hline
  Chapter & Goal & Objective & Research Project & Primary Outcome \\
  \hline
  Chapter~\ref{chapter:case_study}  & RG1 & RO1 & A case study of Web API evolution \cite{sohan2015case}. & A set of challenging problems related to evolving Web APIs. Selected REST API documentation as the problem of interest for this research. \\
  \hline
  Chapter~\ref{chapter:spy_rest} & RG1 and RG2 & RO2 and RO3 & SpyREST: Automated RESTful API documentation using an HTTP proxy server \cite{sohan2015spyrest}. & A list of requirements. A conceptual design using interception to generate REST API documentation to meet the requirements. \\
  \hline
  Chapter~\ref{chapter:demo_paper} & RG2 & RO3 & SpyREST in action: An automated RESTful API documentation tool \cite{sohan2015spyrest_tool}. & A reusable tool that implements the conceptual design and shows how the identified REST API documentation requirements are met.\\
  \hline
  Chapter~\ref{chapter:cisco_study} & RG2 & RO4 & Automated example oriented REST API documentation at Cisco \cite{sohan_cisco}. & Empirical evidence that the proposed technique provides a practical documentation solution for REST API developers.\\
  \hline
  Chapter~\ref{chapter:controlled_study} & RG2 & RO5 & A study of the effectiveness of usage examples in REST API documentation \cite{sohan_vlhcc}. & Empirical evidence that the proposed technique allows REST API client developers to perform API tasks with less effort, higher success and satisfaction ratings.\\
  \hline
\end{tabular}
\label{table:chapter_summary}
\end{table}

@@

In Chapter~\ref{chapter:spy_rest}, I discuss my second research project where I derived a set of REST API documentation requirements and proposed a new technique to meet the requirements \cite{sohan2015spyrest}. Here, I present a list of REST API documentation requirements by studying the literature and applying the findings from Chapter~\ref{chapter:case_study}. I discuss the available REST API documentation techniques and compare against the requirements. I present, \textit{SpyREST}, a novel technique to intercept example REST API calls using an HTTP proxy server to collect and automatically synthesize API traffic to generate and update API documentation to satisfy the identified requirements. The primary advantage of the presented technique is that the documentation is generated from executable code. As a result, a continuously updated REST API documentation with realistic usage examples is achievable with a reduced manual effort. The remaining chapters provide evidence that the proposed technique can satisfy the identified requirements.

I present my third research project, related to RG2, in Chapter~\ref{chapter:demo_paper}, where I present a prototype implementation of the aforementioned technique \cite{sohan2015spyrest_tool}. In this prototype implementation, I discuss the process that transforms a number of example HTTP requests and responses into usable REST API documentation. I also present an example where automated test code is used to generate and maintain a continuously updated REST API documentation using the developed tool.

### 2017-09-24 root.tex
add changes based on feedback from craig + mea

%
%                 Substitute  List of Examiners
%
\begin{signing}{Department of Computer Science}
\signline
Dr. Frank Maurer \\
Supervisor\\
Department of Computer Science \\
University of Calgary \\
Canada\\
\signline
Dr. Mea Wang\\
Supervisory Committee Member\\
Department of Computer Science \\
University of Calgary \\
Canada\\
\signline
Dr. Craig Anslow\\
Supervisory Committee Member\\
School of Eng. and Comp. Science \\
Victoria University of Wellington\\
New Zealand\\
\end{signing}
%
\newpage
\phantomsection
\altchapter{\bf{Abstract}}
API documentation presents both a problem and an opportunity for API usability. REST APIs provide interconnectivity between applications over HTTP. Documentation of a REST API is a key information source for API client developers. Most REST APIs are documented using a manual approach, which can be time consuming and error-prone. REST API developers need to efficiently document their APIs with the qualities that make the API usable. In this research, I focused on the topic of automated REST API documentation to satisfy this need.

In this thesis, I present and evaluate of a novel technique to solve REST API documentation requirements. I present a set of REST API documentation requirements by studying the existing literature and the current industry practice. From this study, I observed that REST APIs evolve frequently, but used a manual or bespoke approach for generating and maintaining their documentation. I present a reusable technique with a prototype implementation to automate the REST API documentation process with executable API usage examples. The technique involves the interception of example REST API calls using an HTTP proxy server to auto-generate an accurate and updated REST API documentation. I present an industrial evaluation of the proposed technique based on a period of eighteen months of production use. From this study, I found that it was feasible to leverage API test code to automatically generate an always-updated REST API documentation for evolving REST APIs with usage examples. To evaluate the impact of usage examples on REST API client developers, I performed a controlled study with experienced REST API developers. From this case study, I found that REST API client developers faced patterns of obstacles using REST APIs that can be reduced by including API usage examples in the API documentation, as suggested in the proposed technique.

@@

\newpage
\phantomsection
\altchapter{\bf{Acknowledgements}}
\begin{singlespace}
I am grateful to Shahana, my wife, for supporting me and the family while I took the time away to work on this research.

>>>>

%
%                 Substitute  List of Examiners
%
\begin{signing}{Department of Computer Science}
\signline
Dr. Frank Maurer \\
Supervisor\\
Department of Computer Science \\
University of Calgary \\
Canada\\
\signline
Dr. Mea Wang\\
Supervisory Committee Member\\
Department of Computer Science \\
University of Calgary \\
Canada\\
\signline
Dr. Craig Anslow\\
Supervisory Committee Member\\
School of Eng. and Comp. Science \\
Victoria University of Wellington\\
New Zealand\\
\end{signing}
%
\newpage
\phantomsection
\altchapter{\bf{Abstract}}
API documentation presents both a problem and an opportunity for API usability. REST APIs provide interconnectivity between applications over HTTP. Documentation of a REST API is a key information source for API client developers. Most REST APIs are documented using a manual approach, which can be time consuming and error-prone. REST API developers need to efficiently document their APIs with qualities that make the API usable. In this research, I focused on the topic of automated REST API documentation to satisfy this need.

In this thesis, I present and evaluate of a novel technique to solve REST API documentation requirements. I present a set of REST API documentation requirements by studying the existing literature and the current industry practice. From this study, I observed that REST APIs evolve frequently, but used a manual or bespoke approach for generating and maintaining their documentation. I present, SpyREST, a reusable technique with a prototype implementation to automate the REST API documentation process with executable API usage examples. The technique involves the interception of example REST API calls using an HTTP proxy server to auto-generate an accurate and updated REST API documentation. I present an industrial evaluation of the proposed technique based on a period of eighteen months of production use. From this study, I found that it was feasible to leverage API test code to automatically generate an always-updated REST API documentation for evolving REST APIs with usage examples. To evaluate the impact of usage examples on REST API client developers, I performed a controlled study with experienced REST API developers. From this case study, I found that REST API client developers faced patterns of obstacles using REST APIs that can be reduced by including API usage examples in the API documentation, as suggested in the proposed technique.

@@

\newpage
\phantomsection
\altchapter{\bf{Acknowledgements}}
\begin{singlespace}
I am grateful to Shahana, my wife, for supporting me and the family while I took an unacceptable amount of the time and attention away from them to work on this research.
